---
title: "R Review"
output: html_notebook
---

For this review, we'll be working with dispatch data from the Lincoln Police Department. You can find the data online here: [http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records](http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records).

Use the Download menu to select a csv download.

Then move the data file you downloaded into your project folder.

### Load packages

Load the tidyverse, janitor and lubridate packages.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

### Load data

Load your data into an object called 'dispatches'.

```{r}
lps_dispatch <- read_csv("LPD_Dispatch_Records.csv")
```

#### Fix column names

Use janitor to make all column names comply with R preferred style - all lowercase and underscores between words.

```{r}
lps_dispatch <- lps_dispatch %>% clean_names()
```

### Analysis questions

#### Datatypes

Look at the documentation for the data. Do all the columns appear to be formatted correctly?

Use lubridate to change the RPT_Date column to ymd format.

```{r}
dispatch <- lps_dispatch %>% mutate(
    CleanDate = ymd(rpt_date)
)

head(dispatch)
```

Find the 13 rows that failed to parse. Why?

How many cases are in the data? How many unique cases?

```{r}
dispatch %>% arrange(rpt_date)
```

There are 13 cases that have a date of "0." Nice, LPD. So, that would be why.

```{r}
dispatch %>%
  summarise(
    case = n()
    )
```
It appears there are 400,480 cases in this csv file. However, those same 13 - that didn't have dates - also don't have cases. 

```{r}
dispatch %>%
  group_by(case) %>%
  summarise(
    case = n()
  )
```
There are 400,459 unique cases. There is probably a better way to get that answer, but it works, I think. 

#### Arranging

What are the oldest and most recent cases in the data?

```{r}
dispatch %>%
  arrange(CleanDate)
```

Oldest: January 1, 2017
Newest: June 6, 2020

#### Filtering

Create a dataframe called 'missing' with just missing persons cases in it.

```{r}
dispatch %>% mutate(
  missing = sum(cfs_legend == "MISSING PERSON")
)
```

Ok, there are 7,921 missing persons in this data.

Use the str_detect function to find all the cases that mention O Street in the address.

```{r}
dispatch %>%
  filter(str_detect(block_address, "O ST", negate = FALSE))
```

#### Counting

Use the count() function to find the number of dispatches to each neighborhood. 

```{r}
dispatch %>%
  group_by(neighbor) %>%
  summarise(
    count = n()
  ) %>% arrange(desc(count))
```

Which neighborhood appears the most in the data?

Well, NA is the most dangerous neighborhood in Lincoln. Wow, who would have thought? 

Do you see any limitations of this data when we are counting by neighborhood?

Clearly, LPD forgot to assign some cases a neighborhood. Actually, they didnt' assign 133,001. That's a big problem. 

#### Mutating

Create a new column called 'year' that includes just the year from date_fixed.

```{r}
dispatch <- dispatch %>%
  mutate(
    year = as.numeric(format(CleanDate,'%Y'))
  )
```

Use the case_when function to create a new categorical variable dividing the rpt_time column into the 24 hours of the day. Be careful to make sure that each time fits into only one value of the new variable.

```{r}
dispatch <- dispatch %>%
  mutate(
    hour_of_day = case_when(
      as.numeric(rpt_time) >= 0 & as.numeric(rpt_time) <= 59 ~ 0,
      as.numeric(rpt_time) >= 100 & as.numeric(rpt_time) <= 159 ~ 1,
      as.numeric(rpt_time) >= 200 & as.numeric(rpt_time) <= 259 ~ 2,
      as.numeric(rpt_time) >= 300 & as.numeric(rpt_time) <= 359 ~ 3,
      as.numeric(rpt_time) >= 400 & as.numeric(rpt_time) <= 459 ~ 4,
      as.numeric(rpt_time) >= 500 & as.numeric(rpt_time) <= 559 ~ 5,
      as.numeric(rpt_time) >= 600 & as.numeric(rpt_time) <= 659 ~ 6,
      as.numeric(rpt_time) >= 700 & as.numeric(rpt_time) <= 759 ~ 7,
      as.numeric(rpt_time) >= 800 & as.numeric(rpt_time) <= 859 ~ 8,
      as.numeric(rpt_time) >= 900 & as.numeric(rpt_time) <= 959 ~ 9,
      as.numeric(rpt_time) >= 1000 & as.numeric(rpt_time) <= 1059 ~ 10,
      as.numeric(rpt_time) >= 1100 & as.numeric(rpt_time) <= 1159 ~ 11,
      as.numeric(rpt_time) >= 1200 & as.numeric(rpt_time) <= 1259 ~ 12,
      as.numeric(rpt_time) >= 1300 & as.numeric(rpt_time) <= 1359 ~ 13,
      as.numeric(rpt_time) >= 1400 & as.numeric(rpt_time) <= 1459 ~ 14,
      as.numeric(rpt_time) >= 1500 & as.numeric(rpt_time) <= 1559 ~ 15,
      as.numeric(rpt_time) >= 1600 & as.numeric(rpt_time) <= 1659 ~ 16,
      as.numeric(rpt_time) >= 1700 & as.numeric(rpt_time) <= 1759 ~ 17,
      as.numeric(rpt_time) >= 1800 & as.numeric(rpt_time) <= 1859 ~ 18,
      as.numeric(rpt_time) >= 1900 & as.numeric(rpt_time) <= 1959 ~ 19,
      as.numeric(rpt_time) >= 2000 & as.numeric(rpt_time) <= 2059 ~ 20,
      as.numeric(rpt_time) >= 2100 & as.numeric(rpt_time) <= 2159 ~ 21,
      as.numeric(rpt_time) >= 2200 & as.numeric(rpt_time) <= 2259 ~ 22,
      as.numeric(rpt_time) >= 2300 & as.numeric(rpt_time) <= 2359 ~ 23,
      )
  )

#case_when(
  #x %% 35 == 0 ~ "fizz buzz",
  #x %% 5 == 0 ~ "fizz",
  #x %% 7 == 0 ~ "buzz",
  #TRUE ~ as.character(x)

```

This is what I was rocking before. A tad more effecient. 

```{r}
dispatch <- dispatch %>%
  mutate(
    hour_of_day = substr(rpt_time, 1, 2)
  )
```

What hour of the day sees the most police action?

```{r}
dispatch %>%
  group_by(hour_of_day) %>%
  summarise(
    total_per_hour = n()
  ) %>% arrange(desc(total_per_hour))
```

It appears that 4-6 p.m. is not the time to be outside in Lincoln. 

#### Grouping and summarizing 

How many drug-related dispatches occurred in each year of the data?

```{r}
dispatch %>% 
  filter(str_detect(cfs_legend, "DRUGS", negate = FALSE)) %>%
  group_by(year, cfs_legend) %>%
  summarise(
    count = n()
  )
```

Create a new column called month. Then using group_by and summarize, find the maximum, minimum and average number of dispatches per month.

```{r}
dispatch <- dispatch %>%
  mutate(
    month = as.numeric(format(CleanDate,'%m'))
  )
```
```{r}
dispatch %>% 
  count(month) %>% 
  arrange(desc(n))
```
```{r}
dispatch %>%
  group_by(month) %>%
  summarise(
    total_per_month = n()
  ) %>% arrange(desc(total_per_month))
```
```{r}
month_dispatch <- dispatch %>%
  group_by(month, year) %>%
  summarise(
    per_month = n()
  )
```
```{r}
summary(month_dispatch)
```

```{r}
month_dispatch %>% ungroup() %>% summarize(min(per_month), max(per_month), mean(per_month))
```

#### Percent change

What was the percent change in total number of dispatches from 2018 to 2019?

```{r}
by_year <- dispatch %>%
  group_by(year) %>%
  summarise(
    yearly_total = n()
  )
```
```{r}
by_year %>%
  mutate(
    change = ((2019 - 2018)/2018)*100
  )
```

That's a poor way to do it, but it got the correct answer. There was a 0.05% change from 2018 to 2019. 

#### Line charts

Using ggplot, create a line chart of the number of cases per month and year. Choose an appropriate color for the line, add a title and labels and a choose a theme.

```{r}
dispatch %>% group_by(CleanDate) %>% summarize(total=n()) %>% ggplot() + geom_line(aes(x=CleanDate, y=total)) +
  labs(x="Date", y="LPD Dispatch Calls", title= "Lincoln Police Department Dispatches Since Jan. 2017", subtitle= "The warmer the weather -- which here is the middle part of the year -- the more dispatch calls.", caption="Source: Lincoln Police Department | Graphic: Will Bauer") + 
  theme_minimal() + theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10)
  ) +
  ggsave("LPDdispatch.png", width = 8, height = 4)
```

What do you observe about the yearly pattern of police dispatches? Why do you suppose that is?

The warmer the weather, the - probably - more crime, and, therefore, more 911 calls. 

#### Column charts

Using ggplot, create a column chart that shows the five most common categories in the cfs_legend column. Apply appropriate design decisions to your chart.

```{r}
type <- dispatch %>%
  group_by(cfs_legend) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  filter(total >= 22193)
```
```{r}
library(scales)
```

```{r}
ggplot(type, aes(x=reorder(cfs_legend, -`total`), weight=`total`)) + geom_bar()
```

Well, that looks like garbage. Let me play around with this. 

```{r}
ggplot(type, aes(x=reorder(cfs_legend, -`total`), weight=`total`)) + 
  geom_bar() + 
  scale_y_continuous(labels=comma) + 
  labs(x="", y="", title="Lincoln Police's Five Most Common Reasons for Dispatch Since 2017", subtitle= "Maybe someone should ask LPD what other means because there is a lot of other since 2017.", caption="Source: Lincoln Police Department | Graphic: Will Bauer") + 
  coord_flip() + 
  theme_minimal() + theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10)
  ) +
  ggsave("5Dispatches.png", width = 8, height = 4)
```
